Light inextensible strings (thread) under tension in the Schwarzschild geometry

### 2021-01-26 string_physics_educator_revision2.tex
typo
start work on issue #30

There does not appear to be a simple analytical solution to
equation~\ref{rdashdash}; but it is a straightforward second order
ODE, amenable to numerical solution and no harder than the equivalent
for null geodesics, a standard homework problem.  Here we consider
only strings exterior to the event horizon, $r>2M$, which have a
well-defined point of closest approach to the singularity.  Because
the system is invariant under~$\phi\longrightarrow-\phi$, the string
must be symmetrical about this point.  If $\r'$ is small, then a
string follows $r''=r-2M$, while tangential null geodesics follow
$r''=r-3M$.  Thus a taut string would {\em look} curved to a static
observer in Schwarzschild coordinates, having a (visual) radius of
curvature of
$r_0^2/M\sqrt{1-2M/r_0}$ (the extra term comes from the fact that
appearance is driven by local $ds$, not Schwarzschild $r$).  At the
photon sphere $r_0=3M$, such a string would appear to have a radius
of~$\sqrt{27}M$.  For a solar mass black hole this would be about
$15\unit{km}$, a readily understandable result.

>>>>

There does not appear to be a simple analytical solution to
equation~\ref{rdashdash}; but it is a straightforward second order
ODE, amenable to numerical solution and no harder than the equivalent
for null geodesics, a standard homework problem.  Here we consider
only strings exterior to the event horizon, $r>2M$, which have a
well-defined point of closest approach to the singularity.  Because
the system is invariant under~$\phi\longrightarrow-\phi$, the string
must be symmetrical about this point.  If $r'$ is small, then a string
follows $r''=r-2M$, while tangential null geodesics follow $r''=r-3M$.
Working to first order, and writing $d$ for the difference in
Schwarzschild $r$ between string and null geodesic, we see that
$d''=M$ and with appropriate boundary conditions
$d=\frac{Mx^2}{2r_0^2}$, where $x=r_0\theta$ is an appropriate
circumferential coordinate.  However, to a static local observer tied
to the string would
$d_\mathrm{local}=Mx^2/\left(2r_0^2\sqrt{1-2M/r_0}\right)$,

$d_\mathrm{local}=d/\sqrt{1-2M/r_0}$,

Thus a taut string would {\em look} curved to a static
observer in Schwarzschild coordinates, having a (visual) radius of
curvature of
$r_0^2/M\sqrt{1-2M/r_0}$ (the extra term comes from the fact that
appearance is driven by local $ds$, not Schwarzschild $r$).  At the
photon sphere $r_0=3M$, such a string would appear to have a radius
of~$\sqrt{27}M$.  For a solar mass black hole this would be about
$15\unit{km}$, a readily understandable result.
